---
output:
  pdf_document: default
  html_document: default
---
# Intro to Dat Science - HW 2
##### Copyright  Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva
# IST687 Section M002

# Professor Anderson

```{r}
# Enter your name here: Pranav Sharma
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

### Reminders of things to practice from last week: 
Assignment arrow  		<- <br>
The combine command 	c( ) <br>
Descriptive statistics 	mean( ) sum( ) max( ) <br>
Arithmetic operators		+ - * /<br>
Boolean operators		>    <    >=    <=   ==   !=  <br>

**This Week:** Explore the **quakes** dataset (which is included in R). Copy the **quakes** dataset into a new dataframe (call it **myQuakes**), so that if you need to start over, you can do so easily (by copying quakes into myQuakes again). Summarize the variables in **myQuakes**. Also explore the structure of the dataframe



```{r}
myQuakes <- quakes
summary(myQuakes)
```

**Step 1:** Explore the earthquake magnitude variable called **mag**  <br>

A.	What is the average magnitude? Use mean() or summary():


```{r}
# mag is a column inside the dataframe myQuakes. We use the $ sign to call the columns. It matches what we got in summary in the last question.
mag<-mean(myQuakes$mag) 
mag
```

B.	What is the magnitude of the largest earthquake? Use max() or summary() and save the result in a variable called **maxQuake**:


```{r}
#using the max function to directly get the value instead of summary
maxQuake<- max(myQuakes$mag)
maxQuake
```

C.	What is the magnitude of the smallest earthquake? Use min() or summary() and save the result in a variable called **minQuake**:


```{r}
#using the min function to directly get the value instead of summary
minQuake <- min(myQuakes$mag)
minQuake
```

D.	Output the **third row** of the dataframe


```{r}
# by writing [3,] we are calling the third row and the blank value signifies calling every column
myQuakes[3,]
```

E. Create a new dataframe, with only the rows where the **magnitude is greater than 4**. How many rows are in that dataframe (use code, do not count by looking at the output) 


```{r}
#newMyquakes represents a dataframe with putting a condition on values of rows in column mag such that if a column value is not greater than 4, those rows are not included.
newMyQuakes <- myQuakes[myQuakes$mag>4,]

#nrow function is used to calculate number of rows in any dataframe
nrow(newMyQuakes)
```

F.	Create a **sorted dataframe** based on magnitude and store it in **quakeSorted1**. Do the sort two different ways, once with arrange() and then with order()

```{r}
#used the mag column to sort as mentioned in the question 
quakeSorted1 <- sort(myQuakes$mag) 
# had to install tidyverse and use the library function to use arrange.
library(tidyverse)
quakeSorted1 <- myQuakes %>% arrange(myQuakes,mag)
#install.packages("tidyverse")

``` 

G.	What are the latitude and longitude of the quake reported by the largest number of stations?  


```{r}
myNewDataframe <- myQuakes[which.max(myQuakes$stations),]
# creating a separate vector for latitude and longitude to show their values
latitude <- myNewDataframe$lat
longitude<- myNewDataframe$long

latitude 
longitude

# we can combine them by c() function and then save them into new vector
lat_long_max <-c(latitude,longitude)
lat_long_max
```

H.	What are the latitude and longitude of the quake reported by the smallest number of stations?  


```{r}
#Filtering values based on condition that number of stations is the minimum
myNewDataframe <- myQuakes[which.min(myQuakes$stations),]
# this time we represented the output in the form of data frame.
lat_long_min_stations <- data.frame(myNewDataframe$lat,myNewDataframe$long)
lat_long_min_stations
```

**Step 3:** Using conditional if statements <br>

I.	 Test if **maxQuake** is greater than 7 (output “yes” or “no”) <br>
**Hint:** Try modifying the following code in R: 


```{r}
if(maxQuake>7){
  print ("Yes")
}else
  print("No")
#if  (100 < 150) "100 is less than 150" else "100 is greater than 150"
```

J.	 Following the same logic, test if **minQuake** is less than 3 (output “yes” or “no”):


```{r}
if(minQuake<3){
  print ("Yes")
}else
  print("No")
```
