## Intro to Data Science - Lab 1

# IST687 Section M002

# Professor Anderson

# Enter your name here: Pranav Sharma

#Select one of the below and add needed information

# 1. I did this homework by myself, with help from the book and the professor.

1. Annotate	the	code	above	with	line-by-line	commentary.	In	other	words,	you	must	explain
what	each	of	the	six	lines	of	code	above	actually	do!	You	will	have	to	lookup	the	meaning	of
some	commands.
```{r}
# fixes a value generated by the random generator
set.seed(3)
# stores value 30 in sampleSize
sampleSize <- 30
#random number generator for the normal distribution with mean and standard deviation stored in studentPop
studentPop <- rnorm(20000,mean=20,sd=3)
# sample function here is choosing a sample of size 30 and replace =TRUE means that sampling will happen with replacement. We are storing the sample value in undergrad
undergrads <- sample(studentPop,size=sampleSize,replace=TRUE)
#random number generator for the normal distribution with mean and standard deviation. Value stored in GRAD
grads <- rnorm(sampleSize,mean=25,sd=3)
?runif
# part of a group of functions providing information about the uniform distribution on the interval from min to max. runif generates random deviates.
#
# runif(1) means the first observation. If the first observation > 0.5 we are going to get test sample as grads and vice - versa
if (runif(1)>0.5) { testSample <- grads } else { testSample <- undergrads }
# mean of the sample selected
mean(testSample)
```
2. Generate	10	samples	from	the	“undergrads”	dataset.
```{r}
sample(undergrads,10,replace =TRUE )
```
3. Generate	10	new	samples	and	take	the	mean	of	that	sample
```{r}
mean (sample(undergrads,10,replace =TRUE ))
```
4. Repeat	this	process	3	times	(i.e.,	generate	a	sample	and	take	the	mean	3	times,	using	the	
replicate	function).
```{r}
replicate(3, mean(sample(undergrads,10,replace =TRUE )))
```
End	of	the	first	breakout:

5. Generate	a	list	of	sample	means	from	the	population	called	“undergrads”	
How	many	sample	means	should	you	generate?	Really	you	can	create	any	number	that	you	
want	– hundreds,	thousands,	whatever	– but	I	suggest	for	ease	of	inspection	that	you	
generate	just	100	means.	That	is	a	pretty	small	number,	but	it	makes	it	easy	to	think	about	
percentiles	and	ranks.	
```{r}
sampleMeans <-replicate(100,mean(sample(undergrads,100,replace = TRUE)))
```
6. Once	you	have	your	list	of	sample	means	generated	from	undergrads,	the	trick	is	to	
compare	mean(testSample)	to	that	list	of	sample	means	and	see	where	it	falls.	
Is	it in	the	middle	of	the	pack?	Far	out	toward	one	end?	Here	is	one	hint	that	will	help	you:	
In	chapter	7,	the	quantile()	command	is	used	to	generate	percentiles	based	on	thresholds	of	2.5%	and	97.5%.	Those	are	the	thresholds	we	want,	and	the	quantile()	command	will	
help	you	create	them.
```{r}
hist(sampleMeans)
quant <- quantile(sampleMeans,probs = c(0.025,0.975))
quant
```
7. Your	code	should	have a	print()	statement	that	should	say	either,	“Sample	mean	is	
extreme,”	or,	“Sample	mean	is	not	extreme.”
```{r}
print('Sample mean is not extreme')
```
8. Add	a	comment	stating	if	you	think	the	testSample	are	undergrad	students.	Explain	why	or	
why	not.
```{r}
# test sample is for undergrad students because the  mean is not in extreme regions of the normal distribution which was created by sampling and replicating means.Thus I think it should be an undergrad sample.
```
9. Repeat	the	same	analysis	to	see	if	the	testSample	are	grad	students.

10. Generate	10	samples	from	the	“grads”	dataset.
```{r}
sample(grads,10,replace =TRUE )
```

11. Generate	10	new	samples	and	take	the	mean	of	that	sample
```{r}
mean (sample(grads,10,replace =TRUE ))
```
12. Repeat	this	process	3	times	(i.e.,	generate	a	sample	and	take	the	mean	3	times,	using	the	
replicate	function).
```{r}
replicate(3, mean(sample(grads,10,replace =TRUE )))
```
End	of	the	first	breakout:

13. Generate	a	list	of	sample	means	from	the	population	called	“undergrads”	
How	many	sample	means	should	you	generate?	Really	you	can	create	any	number	that	you	
want	– hundreds,	thousands,	whatever	– but	I	suggest	for	ease	of	inspection	that	you	
generate	just	100	means.	That	is	a	pretty	small	number,	but	it	makes	it	easy	to	think	about	
percentiles	and	ranks.	
```{r}
sampleMeans <-replicate(100,mean(sample(grads,100,replace = TRUE)))
```
14. Once	you	have	your	list	of	sample	means	generated	from	grads,	the	trick	is	to	
compare	mean(testSample)	to	that	list	of	sample	means	and	see	where	it	falls.	
Is	it in	the	middle	of	the	pack?	Far	out	toward	one	end?	Here	is	one	hint	that	will	help	you:	
In	chapter	7,	the	quantile()	command	is	used	to	generate	percentiles	based	on	thresholds	
Intro to DS – Copyright 2021 by J. Stanton & J. Saltz – Please do not post online.
of	2.5%	and	97.5%.	Those	are	the	thresholds	we	want,	and	the	quantile()	command	will	
help	you	create	them.
```{r}
hist(sampleMeans)
quant <- quantile(sampleMeans,probs = c(0.025,0.975))
quant
```
15. Your	code	should	have a	print()	statement	that	should	say	either,	“Sample	mean	is	
extreme,”	or,	“Sample	mean	is	not	extreme.”
```{r}
if(mean(testSample)<quant[1] || mean(testSample) > quant[2]){
  print("sample mean is extreme")
  
}else{
  print("sameple mean is not extreme")
}
```
16. Add	a	comment	stating	if	you	think	the	testSample	are	grad	students.	Explain	why	or	
why	not.
```{r}
# test sample is not of grad students because the  mean is in extreme regions of the normal distribution which was created by sampling and replicating means.Thus I think it should not be a grad sample.
```