
# Intro to Data Science - lab 10

##### Copyright 2021, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Pranav Sharma
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```
Instructions:	Association	rules mining,	also	known	as	market	basket	analysis,	is	an
unsupervised	data	mining technique	that	discovers	patterns	in	the	form of	if-then
rules. The	technique	is	“unsupervised”	in the	sense	that	there	is	no	prediction	or
classification happening.	We	are	simply	trying	to	find	interesting	patterns.
In	addition	to working	with	“baskets”	of objects,	association	rules	mining	is	good	at
working with	any	kind	of data	that	can	be expressed	as	lists	of	attributes.	For
example,	a	trip	to	Washington	DC	might	consist of	the	following	attributes:	train,
July,	morning	departure,	afternoon	arrival,	Union	Station,	first	class,	express.		
In	these	exercises	we	will work	with	a	built	in	data	set	called	groceries.		Make	sure
to library	the arules	and	arulesViz packages	before	running	the	following:

```{r}
#install.packages("arules")
#install.packages("arulesViz")
library(arules)
library(arulesViz)
data (Groceries) # Load data into memory
myGroc <- Groceries # Make a copy for safety
summary(myGroc) # What is the structure?
```
1. Examine	the	data	structure	that	summary()	reveals.	This	is	called	a	sparse
matrix	and	it	efficiently	stores	a	set	of	market	baskets	along	with	meta-data.
Report	in	a	comment	about	some	of	the	item labels.
```{r}
summary(myGroc)
# There are item transactions with 169 columns and 9835 rows of data. We see whole milk is 2513
```
2. Use	the itemFrequency(myGroc) command	to	generate	a	list	of	item
frequencies.	Save	that	list	in	a	new	data	object. Run str( ) on	the	data	object
and	write	a	comment	describing	what	it	is.	Run sort( ) on	the	data	object	and
save	the	results. Run head( ) and tail( ) on	the	sorted	object	to	show	the
most	and least	frequently	occurring items.	What’s	the	most	frequently
purchased	item?
```{r}
freq_list<-itemFrequency(myGroc)
str(freq_list)
freq_list <- sort(freq_list)
head(freq_list)
tail(freq_list)
# most frequently sold item is whole milk
```
3. Create	a	frequency	plot	with itemFrequencyPlot(myGroc, topN=20)
and	confirm	that	the	plot	shows	the	most	frequently	purchased	item	with	the
left-most	bar.	Write	a	comment	describing	the meaning	of	the	Y-axis.
```{r}
itemFrequencyPlot(myGroc, topN=20)
# as previously stated, whole milk is the most frequently purchased item and it is on the left-most side of the graph
```
4. Create	a	cross	table	with ct	<- crossTable(myGroc,	sort=TRUE).	Examine	the	first
few	rows	and	columns	of	ct	by	using	the	square brackets	subsetting	technique.
For	example,	the	first	two rows	and	first	three	columns would	be	ct[1:2,	1:3].
Write	a	comment	describing	one	of	values.	Write	a	comment	describing	what	is
on	the	diagonal	of	the	matrix.
```{r}
ct<- crossTable(myGroc, sort=TRUE)
View(ct)
ct[1:2,	1:3]
# a diagonal has the raw frequency of that particular item
```
Intro to	DS	–	Copyright	2021	by	J.	Stanton &	J.	Saltz	–	Please	do	not	post	online.	
Week 10	–	Association	Rules	Mining	
5. Run	the	following	analysis:
rules1 <- apriori(myGroc,
parameter=list(supp=0.0008, conf=0.55),
control=list(verbose=F),
appearance=list(default="lhs",rhs=("bottled beer")))
```{r}

rules1 <- apriori(myGroc, parameter=list(supp=0.0008, conf=0.55),control=list(verbose=F),
appearance=list(default="lhs",rhs=("bottled beer")))

```
6. Examine	the	resulting	rule	set	with	inspect(	)	and	make	sense	of	the	results.
There	should	be	four	rules	in	total.
```{r}
inspect(rules1)
# look for the high lift; {soda,liquor,red/blush wine} and {bottled beer} with the highest lift and having high confidence
# second highest lift with {liquor,red/blush wine} and {bottled beer} with confidence 0.90
# highest lift 12.417929
```
7. Adjust	the	support	parameter	to	a	new	value	so	that	you	get	more	rules.
Anywhere	between	10	and	30	rules	would	be	fine.	Examine	the	new	rule	set
with	inspect(	).	Does	your	interpretation	of	the	situation	still	make	sense?
```{r}
rules2<- apriori(myGroc, parameter=list(supp=0.0006, conf=0.45),control=list(verbose=F), appearance=list(default="lhs",rhs=("bottled beer")))
summary(rules2)
inspect(rules2)
# lift of 10.64 with 85 percent confidence : {liquor (appetizer), dishes}=> {bottled beer}
# lift of 11.23 with a 90 percent confidence in {liquour, red/blush wine}     => {bottled beer}
# lift of 12.41 the highest in these set of rules with 100 percent confidence in {soda,liqour,red/blush wine}     => {bottled beer}
```
8. Power User: use	mtcars	to	create	a	new	dataframe	with	factors	(e.g.,	cyl
attribute).	Then	create	an	mpg	column	with	“good”	or	“bad”	(good	MPG	is
above 25).	Convert	the	dataframe	to	a	transactions	dataset	and	then	predict
rules	for having	bad	MPG.
```{r}
data("mtcars")
mtCars<- mtcars
summary(mtCars)
mtAttr <- mtCars[,c(2,7,8,10)]
#mtAttr
mtAttr$cyl <- as.factor(mtAttr$cyl)
mtAttr$qsec <- as.factor(mtAttr$qsec)
mtAttr$cyl <- as.factor(mtAttr$vs)
mtAttr$cyl <- as.factor(mtAttr$gear)
mtAttr$mpg <- 'Bad'
mtAttr$mpg[mtCars$mpg>25] <- 'Good'
mtAttr$mpg <- as.factor(mtAttr$mpg)
mtAttr_trans <- as(mtAttr, "transactions")
```


```{r}
summary(mtAttr_trans)
itemFrequencyPlot(mtAttr_trans,topN=20)
apriori(myGroc, parameter=list(supp=0.0006, conf=0.45),control=list(verbose=F),
appearance=list(default="lhs",rhs=("bottled beer")))
rules_new <- apriori(mtAttr_trans,parameter = list(supp= 0.001, conf = 0.55),control=list(verbose=F),
                     appearance=list(default="lhs",rhs=("mpg=Bad")))
summary(rules_new)

```
